# Introduction

## An Overview of Statistical Learning {-}

>Broadly speaking, supervised statistical learning involves building a statistical model for predicting, or estimating, an output based on one or more inputs. 

>With unsupervised statistical learning, there are inputs but no supervising output; nevertheless we can learn relationships and structure from such data.


### Wage Data {-}

Load the `Wage` data set via the `ISLR2` package:

```{r}
library(ISLR2)
wage <- ISLR2::Wage

head(wage)
```

Load the `tidyverse`, and set my `ggplot2` theme:

```{r message=FALSE}
library(tidyverse)
library(dunnr) # My personal R package
library(patchwork) # For composing plots
extrafont::loadfonts(device = "win", quiet = TRUE)
theme_set(theme_td())
set_geom_fonts()
set_palette()
```

Now attempt to re-create Figure 1.1:

```{r figure1.1, message=FALSE, fig.height=3}
p1 <- wage %>%
  ggplot(aes(x = age, y = wage)) +
  geom_point(color = "lightgrey") +
  geom_smooth(color = "blue")
p2 <- wage %>%
  ggplot(aes(x = year, y = wage)) +
  geom_point(color = "lightgrey") +
  geom_smooth(method = "lm", color = "blue")
p3 <- wage %>%
  # Need to re-label the factor levels
  mutate(education = fct_relabel(education, ~str_extract(., "\\d"))) %>%
  ggplot(aes(x = education, y = wage)) +
  geom_boxplot(aes(fill = education)) +
  theme(legend.position = "none")
p1 + p2 + p3
```

## A Brief History of Statistical Learning {-}

## This Book {-}

>The Elements of Statistical Learning (ESL) by Hastie, Tibshirani, and
Friedman was first published in 2001. Since that time, it has become an
important reference on the fundamentals of statistical machine learning.
Its success derives from its comprehensive and detailed treatment of many
important topics in statistical learning, as well as the fact that (relative to
many upper-level statistics textbooks) it is accessible to a wide audience.

>The purpose of An Introduction to Statistical Learning (ISL) is to facilitate the transition of statistical learning from an academic to a mainstream
field. ISL is not intended to replace ESL, which is a far more comprehensive text both in terms of the number of approaches considered and the
depth to which they are explored.

I may also pull from ESL [@Hastie2009] at times for more detailed treatments of complex topics.

>ISL is based on the following four premises
>1. Many statistical learning methods are relevant and useful in a wide
range of academic and non-academic disciplines, beyond just the statistical sciences.
2. Statistical learning should not be viewed as a series of black boxes.
3. While it is important to know what job is performed by each cog, it
is not necessary to have the skills to construct the machine inside the
box!
4. We presume that the reader is interested in applying statistical learning methods to real-world problems.
