```{r echo=FALSE}
knitr::opts_chunk$set(fig.retina = 2.5, fig.align = "center",
                      fig.width = 6, fig.height = 3)
```

# Resampling Methods

>Resampling methods are an indispensable tool in modern statistics. They
involve repeatedly drawing samples from a training set and refitting a model
of interest on each sample in order to obtain additional information about
the fitted model. For example, in order to estimate the variability of a linear
regression fit, we can repeatedly draw different samples from the training
data, fit a linear regression to each new sample, and then examine the
extent to which the resulting fits differ. Such an approach may allow us to
obtain information that would not be available from fitting the model only
once using the original training sample.

>In this chapter, we discuss two of the most commonly
used resampling methods, *cross-validation* and the *bootstrap*.

Cross-validation is most often used to estimate test error associated with a statistical learning method, whereas the boostrap is most commonly used to provide a measure of accuracy for a given parameter/method.

> The process
of evaluating a model’s performance is known as *model assessment*, whereas 
the process of selecting the proper level of flexibility for a model is known as *model selection*.

## Cross Validation

Sometimes we want to estimate the test error rate using the available training data.
A number of approaches can be used for this.
In this section we consider methods which involve *holding out* a subset of the training data from the fitting process, then applying the model to that hold-out set for model assessment.

### The Validation Set Approach

This simple strategy involves randomly dividing available oberservations into the training and validation set.
The model is fitted on the training set, and used to make predictions on the validation set.
The corresponding metric from the validation set predictions -- usually MSE in the case of a quantitative response -- provides an estimate of the test error rate.
To illustrate this, load the `Auto` data set and R packages:

```{r message=FALSE}
auto <- ISLR2::Auto

library(tidyverse)
library(tidymodels)
library(broom)
library(gt)
library(patchwork) # for composing plots
# Load my R package and set the ggplot theme
library(dunnr)
extrafont::loadfonts(device = "win", quiet = TRUE)
theme_set(theme_td())
set_geom_fonts()
set_palette()
```

Randomly split the data into 50% training and 50% validation, fit on the training set, and compute the MSE on the validation set.
Since I'll be repeating this 10 times to reproduce the figure, make it a function:
^[In the text, they compute the MSE but here I am computing RMSE then squaring it. This is because `yardstick` has a `rmse()` function but not an `mse()` function. If I wanted to, I could define a custom metric [like so](https://yardstick.tidymodels.org/articles/custom-metrics.html).]


```{r}
evaluate_auto_fit <- function(seed) {
  set.seed(seed)
  auto_split <- initial_split(auto, prop = 0.5)
  auto_train <- training(auto_split)
  auto_validation <- testing(auto_split)
  auto_rec <- recipe(mpg ~ horsepower, data = auto_train)
  lm_workflow <- workflow() %>% add_model(linear_reg())
    
  tibble(poly = 1:10) %>%
    mutate(
      lm_rec = map(
        poly, ~ auto_rec %>% step_poly(horsepower, degree = .x)
      ),
      lm_fit = map(
        lm_rec, ~ lm_workflow %>% add_recipe(.x) %>% fit(auto_train)
      ),
      lm_rmse = map_dbl(
        lm_fit,
        ~ augment(.x, new_data = auto_validation) %>%
          rmse(truth = mpg, estimate = .pred) %>%
          pull(.estimate)
      )
    )
}
evaluate_auto_fit(seed = 308)
```

Now reproduce Figure 5.2:

```{r figure 5.2}
d <- map_dfr(1:10, evaluate_auto_fit, .id = "rep")

# Use a 10 color palette from the MetBrewer package
pal <- MetBrewer::met.brewer("Veronese", 10, type = "continuous")

p2 <- d %>%
  #filter(rep == 1) %>%
  ggplot(aes(x = poly, y = lm_rmse^2, fill = rep)) +
  geom_line(aes(color = rep), size = 1) +
  #geom_point(aes(fill = rep), shape = 21,  color = "white", size = 4) +
  expand_limits(y = c(15, 30)) +
  scale_x_continuous(breaks = seq(2, 10, 2)) +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  labs(x = "Degree of polynomial") +
  theme(legend.position = "none")
p1 <- p2 %+% filter(d, rep == 1) +
  geom_point(aes(fill = rep), shape = 21,  color = "white", size = 4) +
  labs(y = "MSE")
p1 | p2
```


## The Bootstrap

## Lab: Cross-Validation and the Bootstrap

## Exercises
