```{r echo=FALSE}
knitr::opts_chunk$set(fig.retina = 2.5, fig.align = "center",
                      fig.width = 6, fig.height = 3,
                      cache.path = "cache/")
```

# Tree-Based Methods

Load the usual packages:

```{r message=FALSE}
library(tidyverse)
library(tidymodels)
library(broom)
library(gt)
library(patchwork)
library(tictoc)

# Load my R package and set the ggplot theme
library(dunnr)
extrafont::loadfonts(device = "win", quiet = TRUE)
theme_set(theme_td_minimal())
set_geom_fonts()
set_palette()
```

Tree-based methods involve *stratifying* or *segmenting* the predictor space into a number of simple regions.
Predictions are typically the mean or mode of the response value for training observations in a region.
Since the set of splitting rules can be summarized in a tree, these types of approaches are known as *decision tree* methods.

>Tree-based methods are simple and useful for interpretation. However,
they typically are not competitive with the best supervised learning aproaches, such as those seen in Chapters 6 and 7, in terms of prediction
accuracy. Hence in this chapter we also introduce bagging, random forests,
boosting, and Bayesian additive regression trees. Each of these approaches
involves producing multiple trees which are then combined to yield a single
consensus prediction. We will see that combining a large number of trees
can often result in dramatic improvements in prediction accuracy, at the
expense of some loss in interpretation.

## The Basics of Decision Trees

### Regression Trees

#### Predicting Baseline Players' Salaries Using Regression Trees

To motivate *regression trees*, we use an example of predicting a baseball player's `salary` based on `years` (number of years played in the major leagues), and `hits` (number of hits made in the previous year).

```{r}
hitters <- ISLR2::Hitters %>% janitor::clean_names()

# As per the text, we remove missing `salary` values and log-transform it
hitters <- hitters %>%
  filter(!is.na(salary)) %>%
  mutate(salary = log(salary))

glimpse(hitters)
```

I'll start with the `trees` library for this example:

```{r}
library(tree)

hitters_tree <- tree(salary ~ years + hits, data = hitters,
                     # In order to limit the tree to just two partitions,
                     #  need to set the `control` option
                     control = tree.control(nrow(hitters), minsize = 100))
```

Use the built-in `plot()` to visualize the tree in Figure 8.1:

```{r figure8.1}
plot(hitters_tree)
text(hitters_tree)
```

To visualize the regions, there is no `broom::tidy()` method for `tree` objects, but I can get the cuts from the `frame$splits` object:

```{r}
hitters_tree$frame$splits
```

```{r}
splits <- hitters_tree$frame$splits %>%
  as_tibble() %>%
  filter(cutleft != "") %>%
  mutate(val = readr::parse_number(cutleft)) %>%
  pull(val)
splits
```

```{r figure8.2}
hitters %>%
  ggplot(aes(x = years, y = hits)) +
  geom_point(color = td_colors$nice$soft_orange) +
  geom_vline(xintercept = splits[1], size = 1, color = "forestgreen") +
  geom_segment(aes(x = splits[1], xend = 25, y = splits[2], yend = splits[2]),
               size = 1, color = "forestgreen") +
  annotate("text", x = 10, y = 50, label = "R[2]", size = 6, parse = TRUE) +
  annotate("text", x = 10, y = 200, label = "R[3]", size = 6, parse = TRUE) +
  annotate("text", x = 2, y = 118, label = "R[1]", size = 6, parse = TRUE) +
  coord_cartesian(xlim = c(0, 25), ylim = c(0, 240)) +
  scale_x_continuous(breaks = c(1, 4.5, 24)) +
  scale_y_continuous(breaks = c(1, 117.5, 238))
```

>
Overall, the tree stratifies
or segments the players into three regions of predictor space: players who
have played for four or fewer years, players who have played for five or more
years and who made fewer than 118 hits last year, and players who have
played for five or more years and who made at least 118 hits last year.

The regions $R_1$, $R_2$, and $R_3$ are known as *terminal nodes* or *leaves* of the tree.
The splits along the way are referred to as *internal nodes* -- the connections between nodes are called *branches*.

A key advantage of a simple decision tree like this is its ease of interpretation:

>
We might interpret the regression tree displayed in Figure 8.1 as follows:
`Years` is the most important factor in determining `Salary`, and players with
less experience earn lower salaries than more experienced players. Given
that a player is less experienced, the number of hits that he made in the
previous year seems to play little role in his salary. But among players who
have been in the major leagues for five or more years, the number of hits
made in the previous year does affect salary, and players who made more
hits last year tend to have higher salaries.


#### Prediction via Stratification of the Feature Space {-}

Roughly speaking, there are two steps in building a regression tree:

1. Divide the predictor space into $J$ distinct non-overlapping regions $R_1, \dots, R_J$.
2. For every observation that falls into the region $R_j$, we make the same prediction: the mean of the respone values of the training observations.

In theory, the regions could be any shape but we choose *boxes* for simplicity.
The goal of step 1 is the finx the boxes that minimize the RSS

$$
\sum_{j=1}^J \sum_{i \in R_j} (y_i - \hat{y}_{R_j})^2.
$$

>
Unfortunately, it is computationally infeasible to consider every
possible partition of the feature space into $J$ boxes. For this reason, we take
a *top-down, greedy* approach that is known as *recursive binary splitting*. The
approach is *top-down* because it begins at the top of the tree (at which point
all observations belong to a single region) and then successively splits the
predictor space; each split is indicated via two new branches further down
on the tree. It is *greedy* because at each step of the tree-building process,
the best split is made at that particular step, rather than looking ahead
and picking a split that will lead to a better tree in some future step.

The first split is defined as a pair of half planes:

$$
R_1 (j,s) = \{X | X_j < s \} \text{ and } R_2 (j,s) = \{ X | X_j \geq s \},
$$

and we seek the values $j$ and $s$ that minimize

$$
\sum_{i: x_i \in R_1 (j,s)} (y_i - \hat{y}_{R_1})^2 +
\sum_{i: x_i \in R_2 (j,s)} (y_i - \hat{y}_{R_2})^2.
$$

Finding the values $j$ and $s$ to minimize the above RSS can be done quite quickly, especially when the number of features $p$ is not too large.

The process is repeated -- looking for the best predictor and best cutpoint in order to split the data further and minimize the RSS within the regions -- expect instead of splitting the entire predictor space, we split one of the previously identified regions.
This results in three regions, which we again look to split.
This continues until a stopping criterion is reached; for instance, we may continue until no region contains more than five observations.

#### Tree Pruning {-}

The above process is likely to overfit the data, leading to poor test set performance.
This is because the resulting tree might be too complex.
A smaller tree with fewer splits/regions might lead to lower variance and better interpretation at the cost of higher bias.
One way to control this is to only choose splits that exceed some high threshold of RSS reduction.
This way is short-sighted however, because a seemingly worthless split early on might be followed by a great split.

>
Therefore, a better strategy is to grow a very large tree $T_0$, and then
*prune* it back in order to obtain a *subtree*. How do we determine the best
way to prune the tree? Intuitively, our goal is to select a subtree that 
leads to the lowest test error rate. Given a subtree, we can estimate its
test error using cross-validation or the validation set approach. However,
estimating the cross-validation error for every possible subtree would be too
cumbersome, since there is an extremely large number of possible subtrees.
Instead, we need a way to select a small set of subtrees for consideration.
>
*Cost complexity pruning* — also known as *weakest link pruning* — gives us cost
a way to do just this. Rather than considering every possible subtree, we
consider a sequence of trees indexed by a nonnegative tuning parameter $\alpha$.
For each value of α there corresponds a subtree $T \subset T0$ such that

$$
\sum^{|T|}_{m=1} \sum_{x_i \in R_m} (y_i - \hat{y}_{R_m})^2 + \alpha |T|
$$
